---
title: "Getting Started with hvtiverse"
author: "John Ehrlinger"
date: today
format: html
vignette: >
  %\VignetteIndexEntry{Getting Started with hvtiverse}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(hvtiverse)
```

## The problem

The HVTI R package family is eleven packages spread across eleven GitHub
repositories. Setting up a machine used to mean eleven `pak::pak()` calls, in
an order you had to know, against repository names that do not always match
the package names — `TemporalHazard` lives in `ehrlinger/temporal_hazard`, and
`hvtiRpropensity` lives in `ehrlinger/hvtiPropensityScores`.

`hvtiverse` reduces that to one call, and adds a way to ask whether you are
current.

## The members

The registry is the package's only piece of data. It maps each package to the
repository it installs from:

```{r members}
hvtiverse_members()
```

## What you have

`hvtiverse_status()` compares what is installed against what is on each
repository's `main` branch:

```{r status, eval = FALSE}
hvtiverse_status()
```

```
hvtiverse - 11 members

  v hvtiRutilities   1.0.10     1.0.10     ok
  v hvtiRdatasets    0.1.1      0.1.1      ok
  ! hvtiRtables      0.9.0      1.0.0      stale
  x hvtiPlotR        -          2.7.6      missing
  ...

i 2 members need updating. Run hvtiverse::hvtiverse_update().
```

The status column takes one of six values:

| Status | Meaning |
|---|---|
| `ok` | installed version equals the latest |
| `stale` | installed version is behind |
| `missing` | not installed |
| `ahead` | you have a local development build |
| `unknown` | GitHub could not be reached |
| `ok-local` | installed; the remote was not consulted |

The object is a plain data frame underneath, so you can use it in a script:

```{r status-df, eval = FALSE}
st <- hvtiverse_status()
st$package[st$status == "stale"]
```

On a machine with no outbound network, skip the remote entirely:

```{r status-local}
hvtiverse_status(remote = FALSE)
```

## Installing and updating

On a new machine, install everything:

```{r install, eval = FALSE}
hvtiverse_install()
```

Thereafter, install only what has fallen behind:

```{r update, eval = FALSE}
hvtiverse_update()
```

### Why it may refuse

`hvtiverse_update()` stops if a member you are about to update is already
attached in the session:

```
Error: Cannot install hvtiPlotR: already loaded in this session.
i Restart R and run this before anything attaches it.
i Pass `force = TRUE` to install anyway (unsafe on Windows).
```

This is not fussiness. A package whose namespace is loaded cannot be safely
overwritten — on Windows the write fails outright and leaves a half-installed
library. Restart R and run the update before you attach anything.

## When something will not install

```{r doctor, eval = FALSE}
hvtiverse_doctor()
```

The doctor adds two environment checks ahead of the status table: your R
version against the family's strictest requirement, and your platform.

`ggRandomForests` and `hvtiRlifetables` both need R 4.4.0 or newer, which is
the single most common reason a member will not install. Note that `hvtiverse`
itself requires only R 4.1.0 — deliberately, so that the diagnostic still runs
on a machine whose R is too old for the members it is reporting on.

## Why GitHub and not CRAN

Every member installs from GitHub `main`, including `ggRandomForests`, which is
also published on CRAN.

The reason is that a finished release is not the same as an available one. In
August 2026, `TemporalHazard` 1.2.0 and `ggRandomForests` 3.5.1 were both
complete but waiting on the CRAN queue, and `hvtiRlifetables` requires
`TemporalHazard (>= 1.2.0)`. Installing from GitHub delivered those releases to
the group immediately.

One consequence is worth knowing: because `ggRandomForests` came from GitHub, a
later `update.packages()` can quietly downgrade it to the CRAN version. If a
version looks wrong, run `hvtiverse_status()`.
