---
title: "Getting Started with hvtiR"
author: "John Ehrlinger"
date: today
format: html
vignette: >
  %\VignetteIndexEntry{Getting Started with hvtiR}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## The problem

The HVTI R package family is eleven packages spread across eleven GitHub
repositories. Setting up a machine used to mean eleven `pak::pak()` calls, in
an order you had to know, against repository names that have not always matched
the package names: `hvtiRpropensity` lived in `ehrlinger/hvtiPropensityScores`
and `TemporalHazard` in `ehrlinger/temporal_hazard` until the 2026-08 renames.

`hvtiR` reduces that to one call, and adds a way to ask whether you are
current.

## The members

The registry is the package's only piece of data. It maps each package to the
repository it installs from:

```{r members}
hvtiR::members()
```

## What you have

`hvtiR::status()` compares what is installed against what is on each
repository's `main` branch. When pak recorded the installed GitHub commit, the
comparison catches new commits even when the package version did not change:

```{r status, eval = FALSE}
hvtiR::status()
```

```
hvtiR - 11 members

  v hvtiRutilities   1.0.10     1.0.10     ok
  v hvtiRdatabuild    0.1.1      0.1.1      ok
  ! hvtiRtables      0.9.0      1.0.0      stale
  x hvtiPlotR        -          2.7.6      missing
  ...

i 2 members need updating. Run hvtiR::update().
```

The status column takes one of six values:

| Status | Meaning |
|---|---|
| `ok` | installed version and recorded commit are current |
| `stale` | installed version or recorded commit is behind |
| `missing` | not installed |
| `ahead` | you have a local development build |
| `unknown` | GitHub could not be reached |
| `ok-local` | installed; the remote was not consulted |

Packages installed without GitHub commit provenance fall back to the version
comparison. The table stays version-focused; commit SHAs are used internally
only to resolve a tie between equal versions.

The object is a plain data frame underneath, so you can use it in a script:

```{r status-df, eval = FALSE}
st <- hvtiR::status()
st$package[st$status == "stale"]
```

On a machine with no outbound network, skip the remote entirely:

```{r status-local, eval = FALSE}
hvtiR::status(remote = FALSE)
```

```
hvtiR - 11 members

  - hvtiRutilities   1.0.10     -          ok-local
  - hvtiRdatabuild    0.1.1      -          ok-local
  - hvtiRtables      1.0.0      -          ok-local
  - hvtiRtemplates   1.0.0      -          ok-local
  - hvtiPlotR        2.7.6      -          ok-local
  - hvtiRlifetables  1.3.0      -          ok-local
  - hvtiRbootstrap   0.2.0      -          ok-local
  - hvtiRpropensity  1.1.0      -          ok-local
  - hvtiBoostmtree   2.0.1      -          ok-local
  - TemporalHazard   1.2.0      -          ok-local
  - ggRandomForests  3.5.1      -          ok-local

i Remote was not consulted; versions shown are installed versions only.
```

## Installing and updating

On a new machine, install everything:

```{r install, eval = FALSE}
hvtiR::install()
```

Thereafter, install only what has fallen behind:

```{r update, eval = FALSE}
hvtiR::update()
```

`hvtiR::update()` also reports `hvtiR`'s own version against GitHub. The
installer is not a member of its own registry, so nothing else would mention
it, and it cannot update itself from inside a running session -- its namespace
is already loaded, which is exactly what the guard below refuses. When the
report says it is behind, reinstall it the way you first installed it:

```{r self, eval = FALSE}
pak::pak("ehrlinger/hvtiR")
```

### Why it may refuse

`hvtiR::update()` stops if a member you are about to update is already
attached in the session:

```
Error: Cannot install hvtiPlotR: already loaded in this session.
i Restart R and run this before anything attaches it.
i Pass `force = TRUE` to install anyway (unsafe on Windows).
```

This is not fussiness. A package whose namespace is loaded cannot be safely
overwritten — on Windows the write fails outright and leaves a half-installed
library. Restart R and run the update before you attach anything.

## Reproducible installs

`hvtiR::install()` resolves every member from GitHub `main`, so it gives you
the newest code rather than a fixed one. Two people running it a week apart can
end up on different commits under the same version number. When an analysis has
to be reproducible, pin the versions with
[renv](https://rstudio.github.io/renv/):

```{r renv, eval = FALSE}
renv::init()      # once per project
hvtiR::install()  # installs into the project library
renv::snapshot()  # records every package, and each member's commit
```

`renv.lock` then names the exact commit behind each member, and
`renv::restore()` rebuilds that library on another machine. This works because
pak records the commit it installed from, which is the field renv reads.

A member installed from a local working copy rather than from GitHub carries no
such commit, so renv has none to record for it.

## When something will not install

```{r doctor, eval = FALSE}
hvtiR::doctor()
```

The doctor reports four things about your environment ahead of the status
table: your R version against the family's strictest requirement, your
platform, whether `pak` is installed to install members with, and whether an
`renv` project is active.

`ggRandomForests` and `hvtiRlifetables` both need R 4.4.0 or newer, which is
the single most common reason a member will not install. Note that `hvtiR`
itself requires only R 4.1.0 — deliberately, so that the diagnostic still runs
on a machine whose R is too old for the members it is reporting on.

## Why GitHub and not CRAN

Every member installs from GitHub `main`, including `ggRandomForests`, which is
also published on CRAN.

Development flows from the family repositories to CRAN, not the other way
around. GitHub `main` is therefore the leading release source, while CRAN is a
downstream republication for members that are published there.

One consequence is worth knowing: because `ggRandomForests` came from GitHub, a
later `update.packages()` can quietly downgrade it to the CRAN version. If a
version looks wrong, run `hvtiR::status()`.

The same replacement can happen to `TemporalHazard`, and there it may not be
harmless. `hvtiRlifetables` requires `TemporalHazard (>= 1.2.0)` in its
`Imports:` field, a requirement R enforces when the namespace loads. If a CRAN
downgrade no longer satisfies that requirement, `hvtiRlifetables` will not
load until `hvtiR::update()` puts the GitHub release back.
