---
title: "The job catalog"
format: html
vignette: >
  %\VignetteIndexEntry{The job catalog}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| include: false
library(hvtiR)
knitr::opts_chunk$set(echo = FALSE)
```

Every job type found in the studies corpus, and the package that owes it.
SAS counts the studies holding a program of each job type, except for `si`
and `mi`, where it counts the studies calling the imputation macro, because
imputation runs inside other jobs. R counts the R jobs already written. Both are distinct counts, so rows do not sum;
`NA` means not yet measured and `0` means measured and none found. This page is
generated from the catalog shipped with the package, so it cannot disagree
with it.

```{r}
j <- jobs()
knitr::kable(as.data.frame(table(
  disposition = j$disposition
)), col.names = c("disposition", "rows"))
```

```{r}
#| results: asis
show <- function(rows) {
  label <- ifelse(is.na(rows$qualifier), rows$prefix,
                  paste0(rows$prefix, "-", rows$qualifier))
  owed <- vapply(rows$replaced_by, function(v) {
    if (!length(v)) "" else paste0("`", paste(v, collapse = "`, `"), "`")
  }, character(1))
  blocked <- ifelse(is.na(rows$blocked_on), "", rows$blocked_on)
  out <- data.frame(
    job = paste0("`", label, "`"),
    name = rows$name,
    SAS = rows$sas_breadth_jobs,
    R = rows$r_jobs,
    disposition = rows$disposition,
    status = rows$status,
    destination = rows$destination,
    `replaced by` = owed,
    `blocked on` = blocked,
    check.names = FALSE
  )
  print(knitr::kable(out, row.names = FALSE))
  cat("\n\n")
}

for (f in unique(j$folder)) {
  cat("## ", f, "\n\n", sep = "")
  show(j[j$folder == f, , drop = FALSE])
}
```
